---
title: "AGSIci - muscle contraction analysis"
author: "Dominika Sulcova"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
```

```{r libraries,message=FALSE,warning=FALSE}
library(tidyverse)
library(readxl)
library(ggeffects)
library(lmerTest)
library(ggplot2)
library(readr)
library(emmeans) 

rm(list=ls())
setwd('C:/Users/uzivatel/UCL/O365G-NOCIONS - dsulcova/AG-SICI/Results')
```

```{r functions,message=FALSE,warning=FALSE}
check_model<-function(lmm,data){
  plot(fitted(lmm),residuals(lmm))
  qqnorm(residuals(lmm))
  qqline(residuals(lmm))
  ggplot(data.frame(cook=cooks.distance(lmm),id = data$subject_ID),
         aes(x=cook,y=id)) +
    geom_point() +
    theme_bw()
}

col_palette = c('firebrick1', 'cornflowerblue', 'gold2', 'forestgreen')
```

# Muscular contraction - single trial analysis

## Import the data

```{r,message=FALSE,warning=FALSE}
data <- read_delim("AGSICI_P1_muscles_trials.csv", col_names = TRUE, delim = ',') %>% 
  select(subject_ID, orientation, intensity, trial, gfp) %>% 
  mutate(subject_ID = factor(subject_ID)) %>% 
  mutate(orientation = factor(orientation, levels = c("along_normal", "along_reversed", "across_normal", "across_reversed"), ordered = TRUE)) %>% 
  mutate(intensity = factor(intensity, ordered = TRUE)) 

summary(data)
```


## Initial data visualization for sanity check

Let's start with a few visualizations of the data.

One dot per observation:

```{r,echo=FALSE,message=FALSE}
data %>%
  ggplot(aes(x = as.character(orientation), y = gfp, 
             color = orientation, group = interaction(as.factor(subject_ID), orientation))) +  
  geom_jitter(height = 0, alpha = .1) +
  scale_color_manual(values = col_palette) +
  theme_classic()
```
Participant average trajectories:

```{r,echo=FALSE,message=FALSE}
data %>%
  group_by(subject_ID, orientation) %>%
  summarise(m = mean(gfp)) %>%
  ggplot(aes(x = as.character(orientation), y = m, color = subject_ID, group = subject_ID))+  
  geom_line()+
  theme_classic()
```

## Model the data

```{r}
mm1 <- lmer(gfp ~ orientation * intensity + (1|subject_ID), data = data)
check_model(mm1, data)
```

The residuals are clearly not gaussian, let's try to transform:

```{r}
mm2 <-lmer(log(gfp) ~ orientation * intensity + (1|subject_ID), data = data)
check_model(mm2, data)
```

let's see what happens if we square root the data

```{r}
mm3 <-lmer(sqrt(gfp) ~ orientation * intensity + (1|subject_ID), data = data)
check_model(mm3, data)
```

That's not better, rather worse... Leet's see, what happens if I remove subject 12

```{r}
mm4 <-lmer(log(gfp) ~ orientation * intensity + (1|subject_ID), data = filter(data, subject_ID != 12))
check_model(mm4, filter(data, subject_ID != 12))
```

## Interpretation of the model

Let's keep all the subjects and log-transform the data --> mm2. 
First let's have a look at the anova output:

```{r}
anova(mm2)
```

```{r,message=F}
data_predicted <- ggpredict(mm2, terms = 'orientation')
plot(data_predicted, connect.lines = TRUE) +
  scale_color_manual(values = col_palette)


g<-plot(data,connect.lines=TRUE,add.data = T)+
  scale_color_manual(values=c(col_palette[1], col_palette[3]))
g$layers[[1]]$aes_params$alpha=.1
g$layers[[1]]$aes_params$size=.2
plot(g)
```

Let's check post hoc comparisons:

```{r}
emm <- emmeans(mm2, specs = ~ orientation * intensity)
emm
pairwise <- pairs(emm)
print(pairwise)
```


```{r}
# pair-wise ad-hoc tests
hypothesis_test(mm2, terms = c('orientation', 'intensity'), test = NULL)
# https://strengejacke.github.io/ggeffects/articles/introduction_comparisons.html
```

# GFP - mean values analysis

## Import the data
```{r,message=FALSE,warning=FALSE}
data_orientation <- read_delim("AGSICI_P1_GFP_mean.csv", col_names = TRUE, delim = ',') %>% 
  group_by(subject, position) %>%
  summarize(orientation = mean(orientation), 
            contraction = mean(contraction)) 
  
data_orientation <- data_orientation %>% 
  rename(subject_ID = subject) %>% 
  mutate(subject_ID = factor(subject_ID)) %>% 
  mutate(position = recode(position, '1' = 'along_normal', '2' = 'along_reversed', '3' = 'across_normal', '4' = 'across_reversed')) %>% 
  mutate(position = factor(position, ordered = TRUE)) %>% 
  rename(gfp = orientation) 

summary(data_orientation)

data_interaction <- read_delim("AGSICI_P1_GFP_mean.csv", col_names = TRUE, delim = ',') %>% 
  select(subject, position, intensity, interaction, contraction) %>% 
  rename(subject_ID = subject) %>% 
  mutate(subject_ID = factor(subject_ID)) %>% 
  mutate(position = recode(position, '1' = 'along_normal', '2' = 'along_reversed', '3' = 'across_normal', '4' = 'across_reversed')) %>% 
  mutate(position = factor(position, ordered = TRUE)) %>% 
  mutate(intensity = recode(intensity, '1' = 100, '2' = 120, '3' = 140)) %>% 
  mutate(intensity = factor(intensity, ordered = TRUE)) %>% 
  rename(gfp = interaction) 

summary(data_interaction)
```

## ORIENTATION: Initial data visualization

```{r,echo=FALSE,message=FALSE}
data_orientation %>%
  ggplot(aes(x = as.character(position), y = gfp, 
             color = position, group = interaction(subject_ID, position))) +  
  geom_jitter(height = 0) +
  scale_color_manual(values = col_palette) +
  theme_classic()
```

Participant average trajectories:

```{r,echo=FALSE,message=FALSE}
data_orientation %>%
  ggplot(aes(x = as.character(position), y = gfp, color = subject_ID, group = subject_ID))+  
  geom_line()+
  theme_classic()
```

## ORIENTATION: Model the data

```{r}
mm1 <- lmer(gfp ~ position + (1|subject_ID), data = data_orientation)
check_model(mm1, data_orientation)
```

Let's try to log-transform:

```{r}
mm2 <-lmer(log(gfp) ~ position + (1|subject_ID), data = data_orientation)
check_model(mm2, data_orientation)
```

A bit better... Now let's prepare a model where we add the contraction as a covariate:

```{r}
mm3 <-lmer(log(gfp) ~ position * contraction + (1|subject_ID), data = data_orientation)
```

## ORIENTATION: Interpretation of the model

Let's keep all the subjects and log-transform the data. 
First let's have a look at the anova output without the covariate:

```{r}
anova(mm2)
```

Now let's see how much influence does the covariate have:

```{r}
anova(mm3)
```
Clearly the size of muscular contraction can explain some of the variance but not all of it, at least not when we average across intensities.

Let's visually check marginal effects: 

```{r,message=F}
data_predicted <- ggpredict(mm2, terms = 'position')
plot(data_predicted, connect.lines = FALSE, add.data = TRUE) 
```

Let's check post hoc comparisons:

```{r}
emm <- emmeans(mm2, specs = ~ position)
pairwise <- pairs(emm)
print(pairwise)
```

## INTERACTION: Initial data visualization

```{r,echo=FALSE,message=FALSE}
data_interaction %>%
  ggplot(aes(x = as.character(position), y = gfp, color = intensity)) +  
  geom_jitter(height = 0, alpha = .25)+
  scale_color_manual(values = col_palette[c(1, 2, 4)], labels = c("100 %rMT", "120 %rMT", "140 %rMT")) +
  geom_jitter(height = 0) +
  theme_classic()
```

## ORIENTATION: Model the data

```{r}
mm1 <- lmer(gfp ~ position * intensity + (1|subject_ID), data = data_interaction)
check_model(mm1, data_interaction)
```

Let's try to log-transform:

```{r}
mm2 <-lmer(log(gfp) ~ position * intensity + (1|subject_ID), data = data_interaction)
check_model(mm2, data_interaction)
```

Much better... Now let's prepare a model where we add the contraction as a covariate:

```{r}
mm3 <-lmer(log(gfp) ~ position * intensity * contraction + (1|subject_ID), data = data_interaction)
```

## ORIENTATION: Interpretation of the model

Let's keep all the subjects and log-transform the data. 
First let's have a look at the anova output without the covariate:

```{r}
anova(mm2)
```

Now let's see how much influence does the covariate have:

```{r}
anova(mm3)
```
Now the size of muscular contraction has a significant impact on the gfp!

Let's visually check marginal effects: 

```{r,message=F}
data_predicted <- ggpredict(mm3, terms = c("position","intensity"))
plot(data_predicted, connect.lines = FALSE, add.data = TRUE) 

```

Let's check post hoc comparisons but let's only contrast different intensities within each coil orientation:

```{r}
emm <- emmeans(mm2, specs = ~ position * intensity)
emm
pairwise <- pairs(emm)
print(pairwise)
```